---
title: "stats170"
author: "shoichiro ueno"
date: "2023-01-26"
output: html_document
---


```{r}
library(Quandl)
library(dygraphs)
Quandl.api_key("") 
```


```{r}
retails_elecronic = Quandl(code="FRED/RSEASN",
            type="ts",
            collapse="monthly",
            meta=TRUE)

retails_hobby = Quandl(code="FRED/RSSGHBMS",
            type="ts",
            collapse="monthly",
            meta=TRUE)

retails_furniture = Quandl(code="FRED/RSFHFS",
            type="ts",
            collapse="monthly",
            meta=TRUE)
```


```{r}
sum(is.na(retails_elecronic))
sum(is.na(retails_hobby))
sum(is.na(retails_furniture))
```

excluding data from 2020 and 2021
```{r}
retails_elecronic = retails_elecronic %>% window(end = c(2019, 12))

retails_hobby = retails_hobby %>% window(end = c(2019, 12))

retails_furniture = retails_furniture %>% window(end = c(2019, 12))
```

Electronic data has a trend and increasing seasonality, so the multiplicative decomposition looks like the first choice.

hobby and furniture data has a trend but no apparent seasonality.
```{r dygraph}
retails_elecronic %>% 
  dygraph() %>% 
  dyRangeSelector

retails_hobby %>% 
  dygraph() %>% 
  dyRangeSelector

retails_furniture %>% 
  dygraph() %>% 
  dyRangeSelector()
```

```{r test data}
retails_electronic_train = retails_elecronic %>% 
  window(end = c(2018, 12))

retails_elecronic_test = retails_elecronic %>% 
  window(start = 2019)

retails_hobby_train = retails_hobby %>% 
  window(end = c(2018, 12))

retails_hobby_test = retails_hobby %>% 
  window(start = 2019)

retails_furniture_train = retails_furniture %>% 
  window(end = c(2018, 12))

retails_furniture_test = retails_furniture %>% 
  window(start = 2019)
```

# decomposition
Comparing the random term between the additive and multiplicative of the electronic data, the one with multiplicative looks more random whereas the additive one still reflects some cyclical pattern.

```{r }
plot(decompose(retails_electronic_train)) 
plot(decompose(retails_electronic_train, type = "mult"))
```

hobby and furniture data do not have a seasonality but have an increasing trend.
```{r}
plot(decompose(retails_hobby_train)) 
plot(decompose(retails_furniture_train)) 
```

Seasonal box-plot also supports that electronic data has seasonality and the hobby and furniture data has no seasonality.
```{r seasonal boxplot}
boxplot(retails_electronic_train ~ cycle(retails_electronic_train))

boxplot(retails_hobby_train ~ cycle(retails_hobby_train))

boxplot(retails_furniture_train ~ cycle(retails_furniture_train))

```

# Acf
To appropriately get the ACF, we need some transformation.
If we apply it to raw data, every lag is significant.
```{r }
acf(retails_electronic_train)
acf(retails_hobby_train)
acf(retails_furniture_train)
```

```{r transformation can help stabilise the variance of time series}
retails_electronic_train_log = log(retails_electronic_train)

retails_electronic_train_quadratic = retails_electronic_train^2

retails_electronic_train_squared = sqrt(retails_electronic_train)

par(mfrow = c(3, 1))
plot(retails_electronic_train_log, main = "log") 
plot(retails_electronic_train_quadratic, main = "quadratic")
plot(retails_electronic_train_squared, main = "squared")
```

ACF include the direct and indirect effect, and PACF only include the direct effect.
```{r diffirence can help stabilise the mean of time series}
log_data = diff(retails_electronic_train_log)
plot(log_data) # now it is a mean of zero and has a constant variance

acf(log_data)
pacf(log_data)
```

# smoothing
Use the raw training dependent variable and fit to it an appropriate exponential smoothing model and forecast with it
```{r model}
# because electronic data has seasonality and trend, we use Seasonal exponential smoothing # plot(retails_electronic_train)

hw_electronic = HoltWinters(retails_electronic_train, 
                            seasonal = 'multiplicative')

plot(hw_electronic)
```

```{r}
preditc_electronic = predict(hw_electronic, n.ahead = 12)
plot(hw_electronic, preditc_electronic)
length(retails_elecronic_test)
```



```{r}
# plot(retails_hobby_train)
# retails_hobby_train has a trend but no seasonality
hw_hobby = HoltWinters(retails_hobby_train, gamma = FALSE)
plot(hw_hobby)
```

```{r}
preditc_hobby = predict(hw_hobby, n.ahead = 12)

plot(hw_hobby, preditc_hobby)
```



```{r}
# plot(retails_furniture_train)
# retails_furniture_train has trend no seasonality

hw_furniture = HoltWinters(retails_furniture_train, gamma = FALSE)
plot(hw_furniture)
```

```{r}
preditc_furniture = predict(hw_furniture, n.ahead = 12)

plot(hw_furniture, preditc_furniture)
```

